---
title: "Lecture 7: Programming"
author: "James Sears* <br> AFRE 891/991 SS 25 <br> Michigan State University"
date: ".small[<br> *Parts of these slides are adapted from [“Data Science for Economists”](https://github.com/uo-ec607/lectures) by Grant McDermott and [“Advanced Data Analytics”](https://github.com/msu-econ-data-analytics/course-materials) by Nick Hagerty.]"

      
output:
  xaringan::moon_reader:
    css: [msu-default.css, msu-metropolis.css, metropolis-fonts]
    lib_dir: libs
    self_contained: true
    nature:
      highlightStyle: github
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
      fig_caption: true

header-includes:
  -\usepackage{amsmath}
---

```{css, echo=FALSE}
# CSS for including pauses in printed PDF output (see bottom of lecture)
@media print {
  .has-continuation {
    display: block !important;
  }
}
.remark-code-line {
  font-size: 95%;
}
.small {
  font-size: 75%;
}
.scroll-output-full {
  height: 90%;
  overflow-y: scroll;
}
.scroll-output-75 {
  height: 75%;
  overflow-y: scroll;
}
```

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(knitr)
library(fontawesome)
knitr::opts_chunk$set(
	fig.align = "center",
	cache = FALSE,
	dpi = 300,
  warning = F,
  message = F,
	fig.height = 5,
	out.width = "80%"
)
```

# Table of Contents

1. [Prologue](#prologue)

2. [If/Else Statements](#ifelse)

3. [For Loops](#for)

4. [Functions](#fun)

5. [Indirection and Name Injection](#indirect)

6. [Vectorization](#vec)

7. [Parallelization](#parallel)


---
class: inverse, middle
name: prologue

# Prologue


---
# Programming

So far in class we've learned how to do a lot of things in R, but we can exponentially increase our data analytics skills (and how quickly we get things done) by learning some .hi-blue[programming].

--


* Write .hi-red[custom functions] to execute specific tasks
  * Scrape all Yellowpages business links for a given search term in hundreds of different cities
--

* .hi-purple[Conditionally] define variables or execute different tasks
  * Create a variable conditional on another variables' values
--

* Perform a repeated task by .hi-pink[looping over values]
  * Create a set of state-level dummy variables from state FIPS codes
--

* Run tasks efficiently .hi-medgrn[in parallel]
  * Calculate parcel or farm-level measures of precipitation and temperature


---
# Programming

Packages we'll use today:

```{r}
pacman::p_load(dslabs, tidyverse, furrr, tictoc, future, progressr)
```

--

And let's load in the `murders` data from the `dslabs` package:

```{r}
data(murders)
```
---
class: inverse, middle
name: ifelse

# If/Else Statements

---
# If/Else Statements

If/else statements are a type of .hi-medgrn[conditional expression].
  * Check to see if a logical condition is True

--
  
  * If True, do a thing
--
  
  * If False:
    * Do a different thing,
    * Do nothing, or
    * Check *another* condition, do a thing if True, etc.

---

# If/Else Statements

For example: print the reciprocal of `a`, unless `a` is 0.

```{r}
a = 0

if(a != 0) {
  
  print(1 / a)
  
} else {
  
  print("Reciprocal does not exist.")
  
}
```

--

Statements like this are used for .hi-blue[control flow] of your code.
* Used all the time in software development
* Used occasionally in data analysis, more often in custom functions and packages.


---

# If/Else Statements

You can also link together multiple condition with `else if`s.
```{r}
if(a > 0) {
  
  print("a is Positive")
  
} else if (a < 0){
    
  print("a is Negative")

} else {
  
  print("a is Zero")
}
  

```

---

# If/Else Statements
A related function that you *will* use all the time in data analysis: `ifelse`. 

--

.center[
syntax: `ifelse(CONDITION, ACTION_IF_TRUE, ACTION_IF_FALSE)`
]

  * `CONDITION`: a logical condition
  *`ACTION_IF_TRUE`: what to do if the condition is true 
  *`ACTION_IF_FALSE`: what to do if the condition is false

--

For example:
```{r}
a = 0
ifelse(a > 0, 1/a, NA)
```

---

# If/Else Statements

.center[
syntax: `ifelse(CONDITION, ACTION_IF_TRUE, ACTION_IF_FALSE)`
]



`ifelse` is particularly useful because it is .hi-purple[vectorized] and can be applied over  .hi-purple[a vector of elements all at once] 

--

For example, to change negative numbers to missing:
```{r}
b = c(0, 1, 2, -3, 4)
ifelse(b < 0, NA, b)
```


---

# If/Else Statements

.center[
syntax: `ifelse(CONDITION, ACTION_IF_TRUE, ACTION_IF_FALSE)`
]



`ifelse` is particularly useful because it is .hi-purple[vectorized] and can be applied over  .hi-purple[a vector of elements all at once] 



Or for adding a conditional variable - for example, whether or not a state is Michigan

```{r}
murders <- murders %>% mutate(
    is_michigan = ifelse(state == "Michigan", "Is Michigan", "Is Not Michigan") #<<
                  )
murders[c(1, 23:26),]
```


---

# if_else

Note that tidyverse has its *own, expanded* version of this function: `if_else()`. Benefits include

--

  * Can use the optional argument `missing` to handle NA values of the condition
  
```{r}
x <- c(-2:0, NA, 2)
ifelse(x < 0, "Negative", "Positive")
if_else(x < 0, "Negative", "Positive", "Missing")
```

---

# if_else

Note that tidyverse has its *own, expanded* version of this function: `if_else()`. Benefits include


  * Preserving data types!
  
```{r}
x <- factor(sample(letters[1:5], 10, replace = TRUE))
ifelse(x %in% c("a", "b", "c"), x, NA)
if_else(x %in% c("a", "b", "c"), x, NA)
```

---
# case_when()

While it's technically possible to use nested ifelses, friends don't let friends nest ifelses.

--

Instead, use .hi-slate[dplyr's] `case_when()`

```{r}
x <- 1:10
## dplyr::case_when()
case_when(
  x <= 3 ~ "small",
  x <= 7 ~ "medium",
  TRUE ~ "big"      # Default value
  )
```
---
# case_when()

Works great within `mutate()` as well!

```{r}
murders <- murders %>% mutate(
    my_opinion = case_when(
      state == "Michigan" ~ "Great State",
      state %in% c("California", "Hawaii") ~ "Also Solid State",
      state == "Missouri" ~ "More like Misery am I right",
      TRUE ~ "A State")
    )
murders[c(1, 5, 12, 23, 26, 38),c(1,7)]
```

---
class: inverse, middle
name: for

# For Loops

---
# Abstraction

Often you will have tasks where you find yourself copying and pasting your code to do the same thing $n$ times, with only minor tweaks each time.

Q: What's wrong with that?

--

A: It's:
--

  * Annoying (especially if $n$ is large)
--
  
  * Hard to change later if needed
--
  
  * Prone to errors/bugs

Instead, you can .hi-medgrn[abstract] your code: define it once, and run it multiple times. The rest of this lecture covers tools for abstraction in different situations.

A good rule to aim for is to .hi-blue[never copy-and-paste more than twice.] If you're pasting more than that, abstract it instead!

---
# Abstraction Methods

There are several different methods for code abstraction that we'll go over:
  
  1. .hi-blue[For loops:] when you want to repeat the same code for .hi-blue[different values of a variable or vector]
--

  1. .hi-medgrn[Functions:] when you want to repeat the same code for potentially .hi-purple[different values of all arguments/variables] or with .hi-medgrn[different settings/samples]
--

  1. .hi-red[Vectorization and Functionals:] when you want to .hi-red[repeat a function] over .hi-purple[different values of arguments]
  
---
# For Loops

The .hi-blue[for loop] is a simple tool for .hi-blue[iteration]


```
for (INDEX in RANGE){
  action(INDEX)
}
```


  * `INDEX` the name of the index you want to use (often `i` but can be anything)
    
    * Often `i` but can be literally anything R-legal
--
  
  * `RANGE` the vector of values to iterate over 
    * Can be numbers, characters, or objects
--

  * `action` the stuff to do for each index value in the specified range
  
---
# For Loops

.hi-medgrn[Good use] for a for loop: repeat the same thing with minor tweaks.

```{r}
for (i in 1:6){
  print(paste0("It is ", i, " O'Clock."))
}
```


---
# For Loops

You can also combine for loops with if-else:

```{r}
for (i in c("Indiana", "Michigan", "Colorado")){
  if (i == "Michigan"){
      print("This is Michigan")
  } else {
      print("This is not Michigan")
  }
}
```


---
# For Loops

Suppose you wanted to calculate the mean of the numeric variables in `murders` and the murder rate. We could manually type and copy-paste:

```{r}
murders <- mutate(murders, rate = total/population * 1e5)
mean(murders$total)
mean(murders$population)
mean(murders$rate)
```

---
# For Loops

Or we could avoid copy-past errors and use a for loop:

```{r}
for (var in c("total", "population", "rate")){
print(mean(murders[[var]]))
}
```

---
# For Loops

We an also loop over an .hi-blue[object in memory:]

```{r}
numeric_col <- c("total", "population", "rate")
for (var in numeric_col){
print(mean(murders[[var]]))
}
```
---
# For Loops

Or .hi-medgrn[assign output to memory] too
```{r}
numeric_col <- c("total", "population", "rate")
means <- vector() # initiate an empty vector
for (var in numeric_col){
means[[var]] <- mean(murders[[var]])
}
```


--

</br>

There is one technical problem with this code. The vector storing the output .hi-purple["grows" at each iteration], which can make the loop .hi-purple[very slow].

---
# For Loops

.hi-blue[Better:] give your empty vector the .hi-blue[right length] *before* starting.

```{r}
means <- vector("numeric", length = length(numeric_col)) # initiate an empty vector with same length
for (i in 1:length(numeric_col)){
  col_num <- which(colnames(murders) == numeric_col[i])
  means[[i]] <- mean(murders[[col_num]])
}
```

---
# For Loops: Caveat

For-loops are actually .hi-blue[discouraged in R programming].

  * We're covering them because the concepts are foundational.

  * But R has nicer ways to iterate, called .hi-medgrn[vectorization].
    * Often *exponentially* faster
    
  * To do proper vectorization, we first need to know how to .hi-purple[write functions].

---
class: inverse, middle
name: fun

# Functions

---
# Functions

We've already seen a .hi-medgrn[multitude of functions] in R
  
  * pre-packaged with base R
  * loaded by different packages (e.g. `dplyr::mutate()`)
  
Regardless of where they come from, the all follow the same basic syntax:

.hi-center[
`function_name(ARGUMENTS)`
]


---
# Custom Functions

While we will often use pre-made functions, you can - and should! - write your own functions too. This is easy to do with the generic **`function()`** function.<sup>1</sup>

If you only have a short function, you can write it all on a .hi-medgrn[single line:]
```{r, eval = F}
function(ARGUMENTS) OPERATIONS
```


.footnote[<sup>1.</sup> Yes, it's a function that let's you write functions. Very meta.] 



---
# Custom Functions

Oftentimes we want our function code to span .hi-pink[multiple lines]. In this case we can use brackets around the function's operations:

```{r, eval = F}
function(ARGUMENTS) {
  OPERATIONS
  return(VALUE)
}
```




---
# Custom Functions

Rather than write .hi-medgrn[anonymous] functions, .hi-blue[name your functions] to assign them to memory and reuse them throughout your file:

```{r, eval = F}
my_func <- function(ARGUMENTS) { #<<
    OPERATIONS
    return(VALUE)
  }
```

--

Try to give your functions short, pithy names that are 
  
  * Informative to you
  * Clear to anyone else who might read the code
  
---
# Building Custom Functions

Let's start with a basic function: calculate a .hi-medgrn[number's square].<sup>2</sup>

```{r}
square <-      # function name
  function(x){ # the arguments of our function (here just one)
  x^2          # the operation(s) that our function performs
  }
```


Testing:
```{r}
square(4)
```


.footnote[<sup>2</sup> I want to note that this .hi-blue[isn't a useful function]. R's arithmetic functions already handle vectorised exponentiation and do so very efficiently.]


  
---

# Specifying Return Values

We can .hi-blue[specify return values] with `return()`
  * Helpful when our function performs a bunch of .hi-medgrn[intermediate steps]

```{r}
square_plus3 <- function(x){ 
  x_sq <- x^2 # assign squared value as intermediate object
  x_sq_pl_3 <- x_sq + 3 
 return(x_sq_pl_3) #<<
  }
```

 
---

# Specifying Return Values

Testing:

```{r}
square_plus3(3)
```

Note that the intermediate objects .hi-green[don't stay in memory] - they're automatically removed as soon as the function is done running.

--

If we left out the `return()`, the function will return .hi-pruple[the result of the very last operation]
  * Unless you assigned it to memory, in which case you get nothing
---

# Specifying Return Values

If we want to return .hi-purple[multiple objects] from our function, we need to either

.hi-medgrn[1\. Return a List]

  
```{r}
square_ret_list <- function(x){ 
  x_sq <- x^2 # assign squared value as intermediate object 
  res <- list(value = x, val_squared = x_sq) #<<
  return(res) 
  }

square_ret_list(3)
```
---

# Specifying Return Values

If we want to return .hi-purple[multiple objects] from our function, we need to either

.hi-blue[2\. Build a data frame] (a tidy solution!)

  
```{r}
square_ret_df <- function(x){ 
  x_sq <- x^2 # assign squared value as intermediate object 
  res <- data.frame(value = x, val_squared = x_sq) #<<
  return(res) 
  }

square_ret_df(3)
```

---

# Default Argument Values

We can also assign .hi-medgrn[default argument values]
  * Allows for all/any arguments to be optional
  * Uses the supplied value when supplied
  * Uses provided default value when not
  
--

Suppose we wanted to expand our function to do any exponent and not just squares:

```{r}
raise_power <- function(x = 2, power = 2){
  res <- data.frame(
      value = x,
      power = power,
      value_raised = x^power
  )
  return(res)
}
```



---

# Default Argument Values

Setting default values doesn't affect typical function usage:

```{r}
raise_power(x = 5, power = 3) # uses specified values
```

--

But now any argument that we omit will .hi-green[use the default values] and the function will run:

```{r}
raise_power() # uses default values of x and power = 2
```

---

# Default Argument Values

Setting default values doesn't affect typical function usage:

```{r}
raise_power(x = 5, power = 3) # uses specified values
```

Without supplying argument values, our previous function wouldn't have worked:

```{r, error = T}
square()
```


---
class: inverse, middle
name: indirect

# Indirection and Name Injection

---

# Indirection

A common use-case for custom functions is .hi-purple[iterating over variables]

  * Repeat a cleaning task over multiple variables in a data frame
  * Run analysis with a different dependent variable

--

For example, let's go back to our `square` function. By default it applies over an entire vector:


```{r}
square(murders$rate)
```

---

# Indirection

We could use it *within* a mutate if we want a new column in our data frame: 
```{r}
murders <- murders %>% 
  mutate(rate_sq = square(rate))
select(murders, starts_with("rate")) %>% head()
```
But doing this for a lot of variables would require a lot of typing (and wouldn't vectorize over multiple variables well)


---

# Indirection

What we might want to do is modify our function to  use .hi-purple[variable names and the dataframe] as the arguments to directly add a new variable:

```{r}
square_df <- function(var, # variable to square
                      df){ # data frame to square variables in
  df <- mutate(df, 
               newvar = var * var)
  return(df)
}
```


---

# Indirection

However, if we try and use this function on the `rate` variable in `murders` with a string, we get an error:

```{r, error = T}
square_df(
  var = "rate", 
  df = murders)
```


---

# Indirection

We get a similar error if we give the variable argument as a .hi-blue[data-variable] 
  * .hi-blue[data-variable]: a "statistical" variable that lives .hi-blue[in a data frame]
    * i.e. an object in the data frame (how we've been using variables in tidy verbs so far)
    
    
```{r, error = T}
square_df(
  var = rate, 
  df = murders)
```


---

# Indirection

This is an issue of .hi-purple[indirection], which occurs in cases like this
  - Want to interpret the argument as an .hi-medgrn[environment-variable] rather than as a .hi-blue[data-variable].
  - .hi-medgrn[env-variable]: "programming" variable/object that lives in your environment (i.e. data frame created with `<-`)
    
<br>

Fortunately, there are a couple programmatic ways around this.

---

# Indirection

.hi-purple[Solution A:] provide the argument as a .hi-blue[data-variable], then 
  1. .hi-purple[embrace] the argument with double braces `{{ arg }}` within function operations 

--

```{r}
square_embr <- function(var, # data-var rather than a string 
                      df){ 
  df <- mutate(df, 
               newvar = {{ var }} * {{ var }} ) #<<
  return(df)
}
square_embr(rate, murders) %>% select(rate, newvar) %>% head()
```

---

# Indirection

.hi-purple[Solution B:] provide the argument as a .hi-blue[data-variable], and 
  1. .hi-red[defuse] the string with `enquo()`
  
  1. .hi-dkgrn[inject] the defused string back in operations with `!!defused_string`

Let's work through the steps involved.

---
# Defusing and Injecting

.hi-red[Defusing] an argument with  .hi-red[delays its evaluation].
  * Use `enquo()` for defusing a single argmuent, `enquos()` for multiple simultaneously
  
For example, returning a defused variable highlights the different handling: 

```{r}
def_fun <- function(df, var){
  # defuse 
  var <- enquo(var) #<<
  return(var)
}
def_fun(murders, rate)
```

---
# Defusing and Injecting

When we're ready to do something with the data-variable, we .hi-dkgrn[inject] it back in
  * Add `!!` in front of the data-variable in the operation
  

```{r}
def_inj_fun <- function(df, var){
  # defuse 
  var <- enquo(var)
  res <- pull(df, !!var) #<<
  return(res)
}
def_inj_fun(murders, rate) %>% head()
```

---
# Defusing and Injecting

.less-left[
Using the two in our square function involves
  1. Defuse the data-variable argument `var`
  1. Inject it in the squaring operation with `!!`
]
.more-right.font90[
```{r}
square_def <- function(var, # data-var rather than a string 
                      df){ 
  var <- enquo(var) # defuse the string #<< 
  
  df <- mutate(df, 
               newvar = !!var * !!var # square the defused string #<<
               ) 
  return(df)
}
square_def(rate, murders) %>% select(rate, newvar) %>% head()
```
]


---

# Defuse + Inject

.hi-purple[Solution C:] defuse the argument with `ensym()`
  * Allows for supplying the argument as either a .hi-pink[character string] or a .hi-blue[data variable]

```{r}
square_ensym <- function(var, df){ 

  df <- mutate(df, 
               newvar = !!ensym(var) * !!ensym(var) # square the defused string #<<
               ) 
  return(df)
}
```
---

# Defuse + Inject

.hi-purple[Solution C:] defuse the argument with `ensym()`
  * Allows for supplying the argument as either a .hi-pink[character string] or a .hi-blue[data variable]

```{r}
square_ensym("rate", murders) %>% select(rate, newvar) %>% head(3)
square_ensym(rate, murders) %>% select(rate, newvar) %>% head(3)
```

---

# ensym vs. enquo vs. {{ }}


`ensym()` and `enquo()` are two of the [tidy defusal operators](https://rlang.r-lib.org/reference/defusing-advanced.html). But what's the difference?
--

  * `ensym()` en"symbolizes" the argument and turns it into a .hi-pink[simple symbol].
--

  * `enquo()` converts the argument to a .hi-red[quosure]

--

When combined with `!!` they allow for pretty advanced defusal operations.

--

.hi-slate[Punchline:] in most cases, .hi-purple[embracing] with `{{ }}` .hi-purple[works in most cases] and is .hi-purple[simpler]

---

# Name Injection

We can combine defusing or embracing with .hi-purple[name injection] to customize our variable names.
  * i.e. call the new squared rate variable `rate_sq` rather than `newvar`

Often we want to programmatically create new variable names based either on
 1. A .hi-medgrn[supplied character string] as a function argument, or
 1. .hi-blue[Modify the data-variable's name] directly in the function
 
 
---

# Name Injection


.hi-medgrn[Approach 1:] use .hi-pink[glue syntax] and .hi-blue[supply the new name as a third argument]:
  * argument `newname` the new variable's name (as a character string)
  
  <br>
  <br>
  <br>
  
```{r, eval = F}
square_inj_1 <- function(var, df,
                       newname){ # new variable name to use #<<
  

  
  return(df)
}
```


---

# Name Injection


.hi-medgrn[Approach 1:] use .hi-pink[glue syntax] and .hi-blue[supply the new name as a third argument]:
  * argument `newname` the new variable's name (as a character string)
  * Glue syntax with `"{newname}"`
  
  <br>  
  <br>
  
```{r, eval = F}
square_inj_1 <- function(var, df,
                       newname){ # new variable name to use 
  
  df <- mutate(df, 
               "{newname}"  #<<
  return(df)
}
```
---

# Name Injection


.hi-medgrn[Approach 1:] use .hi-pink[glue syntax] and .hi-blue[supply the new name as a third argument]:
  * argument `newname` the new variable's name (as a character string)
  * Glue syntax with `"{newname}"`
  * Programmatic assignment operator `:=` instead of `=`

  <br>
  
```{r, eval = F}
square_inj_1 <- function(var, df,
                       newname){ # new variable name to use 
  
  df <- mutate(df, 
               "{newname}" :=  #<<
  return(df)
}
```

---

# Name Injection


.hi-medgrn[Approach 1:] use .hi-pink[glue syntax] and .hi-blue[supply the new name as a third argument]:
  * argument `newname` the new variable's name (as a character string)
  * Glue syntax with `"{newname}"`
  * Programmatic assignment operator `:=` instead of `=`
  * quasi-quotes `{{ }}` evaluate the argument contents in the context of the data frame
  
```{r}
square_inj_1 <- function(var, df,
                       newname){ # new variable name to use 
  
  df <- mutate(df, 
               "{newname}" := {{ var }} * {{ var }} ) #<<
  return(df)
}
```

---

# Name Injection


.hi-medgrn[Approach 1:] use .hi-pink[glue syntax] and .hi-blue[supply the new name as a third argument]:
  * argument `newname` the new variable's name (as a character string)
  * Glue syntax with `"{newname}"`
  * Programmatic assignment operator `:=` instead of `=`
  * quasi-quotes `{{ }}` evaluate the contents in the context of the data frame
  
```{r}
square_inj_1(rate, murders, "rate_sq") %>% select(rate, rate_sq) %>% head()
```
---

# Name Injection


.hi-medgrn[Approach 1] works with `ensym()` too 
  
```{r}
square_inj_1b <- function(var, df,
                       newname){ # new variable name to use #<<
  
  df <- mutate(df, 
               "{newname}" := !!ensym(var) * !!ensym(var) ) #<<
  return(df)
}
square_inj_1b("rate", murders, "rate_squared") %>% select(rate, rate_squared) %>% head()
```



---

# Name Injection

.hi-blue[Approach 2A:] use .hi-pink[glue syntax] and .hi-blue[create the name from the data-variable]:
  * `expr()` "defuses" the supplied expression
    * Converts the data-variable (i.e. `rate`) to a name
  * Glue syntax with `"{newname}"`
  * Programmatic assignment operator `:=` instead of `=`
  
  
```{r}
square_inj_2a <- function(var, df){ 
    new_var <- expr(rate) %>% paste0("_sq") # create new variable name internally #<<
    
  df <- mutate(df, 
               "{new_var}" := {{ var }} * {{ var }} ) # glue syntax to assign new name #<<
  return(df)
}
square_inj_2a(rate, murders) %>% select(rate, rate_sq) %>% head()
```



---

# Name Injection

.hi-blue[Approach 2B:] use .hi-pink[glue syntax] and .hi-red[embracing]:
  * Glue syntax with `"{{newname}}_sq"` (no intermediate name object)
  * Programmatic assignment operator `:=` instead of `=`
  
```{r}
square_inj_2b <- function(var, df){ 

  df <- mutate(df, 
               "{{ var }}_sq" := {{ var }} * {{ var }} ) # Glue syntax without intermediate object #<<
  return(df)
}
square_inj_2b(rate, murders) %>% select(rate, rate_sq) %>% head()
```



---

# Name Injection

.hi-slate[Approach 2C:] you guessed it, `ensym()` still works, allowing the `var` argument to be *either* a data-variable or string.
```{r}
square_inj_2c <- function(var, df){ 

  df <- mutate(df, 
               "{{ var }}_sq" := !!ensym(var) * !!ensym(var) ) # Glue syntax without intermediate object #<<
  return(df)
}
square_inj_2c("rate", murders) %>% select(rate, rate_sq) %>% head()
```

---
class: inverse, middle
name: vec

# Vectorization


---

# Vectorization

Where the real benefits of custom functions, indirection, and name injection come in are with .hi-medgrn[vectorization] and .hi-purple[functionals].



These approaches give a new way to repeatedly iterate a function over a vector of argument values.

<br> 

--

Two main approaches:
  1. .hi-pink[apply] family
    - `apply()`, `lapply()`, `sapply()`, `mapply()`
  1. Tidy.hi-blue[map]  .hi-green[list] functions in .hi-slate[purrr]
    - `map()` and `map2()` with `list_c()`, `list_rbind()`, `list_cbind()`
      - Recently superseded the `map_dfr()`, `map_dfc()` functions


---

# apply Family

The base R .hi-pink[apply] family  gives methods for iterating a function over a vector of arguments depending on the format and type of output we want

| Function         | Description                         | Output Type |
|------------------|-------------------------------------|-------------|
| `lapply(X, FUN)` | apply `FUN` to every element of `X` | list        |
| `sapply(X, FUN)` | apply `FUN` to every element of `X` | vector, matrix, or array |
| `vapply(X, FUN)` | `sapply` with specified output types | vector or array |
| `mapply(FUN, ARG1, ARG2, ...)` | multivariate version of `sapply` | list|
| `apply(X, MARGIN, FUN)` | apply `FUN` to every element of `X` over dimension `MARGIN` | vector, matrix, array, or list|

---
# Apply

Suppose you wanted to standardize all the numeric variables in the `murder` data.

You might write a function like this:

```{r}
calculate_z = function(x) {
  z = (x - mean(x)) / sd(x)
  return(z)
}
```


---
# apply Functions

However, applying it over all the numeric variables at once leads to this:

```{r, error = T}
numeric_cols = c("total", "population", "rate")
murder_numbers = murders[numeric_cols]

calculate_z(murder_numbers)
```

This is an example of a function that .hi-slate[isn't vectorized.]


---
# apply Functions

While we could put our function into a for loop, a more efficient/legible approach would use `sapply`<sup>3</sup>:

.center[
`sapply(X, FUN)`
]
```{r}
sapply(murder_numbers, calculate_z) %>% head()
```

.footnote[<sup>3.</sup> `sapply` is an example of a .hi-dkgrn[functional:] a function that takes another function as an argument.]

---
# map and list_ Functions

The tidy alternative to the apply functions are the `map_` family in .hi-slate[purrr]
  * Work a lot like the `apply_` functions, but with tidyverse syntax
  * Combine with `list_` functions to convert to a vector or dataframe

  
| Function         | Description                         | Output Type |
|------------------|-------------------------------------|-------------|
| `map(X, FUN)` |apply `FUN` to every element of `X` | list |
| `map2(X1, X2, FUN)` | apply `FUN` to every element of `X1` and `X2` | list |
| `list_c()` | combine list elements into a vector | vector       |
| `list_rbind()` | combines elements into a data frame row-wise | data frame |
| `list_cbind()` | combines elements into a data frame column-wise | data frame |


---
# map()

Just like with `sapply()` we can iterate our `calculate_z()` over all numeric variables:

.center[
`map(X, FUN)`
]
```{r}
z_map <- map(murder_numbers, calculate_z)
class(z_map)
z_map
```

---

# list_

The `list_` functions provide a convenient way to convert `map()` output directly to a dataframe:
  * Loop our `square_inj_2c()` function over all three numeric variables
  * Combine each of the dataframes
```{r}
map_sq <- map(
  c("total", "population", "rate"), # first argument: variable names
  square_inj_2c, # function to iterate over
  df = murders # additional static arguments
) %>%
  list_cbind(name_repair = "unique") # account for duplicated names
class(map_sq)
colnames(map_sq)
```

---
class: inverse, middle
name: parallel

# Parallelization


---

# Parallelization

One distinct advantage of R over Stata is the ability to .hi-medgrn[run code in parallel]
  * i.e. split a repeated task across multiple CPU cores simultaneously
  * Useful in any situation where we would use `map()` - i.e. bootstrapping, extracting parcel-level raster information
    
--

.pull-left[
.center.hi-medgrn[Stata]
  * SE: runs in "serial" on one core
  * MP Student: 4 core ($375/yr)
  * MP 8 Core: $655/yr
]

--

.pull-right[
.center[.hi-blue[R] and .hi-slate[furrr]]
  * `future_map` functions work exactly like .hi-slate[purrr's] `map()`
  * Run across as many cores as your system has
  * See progress with .hi-slate[progressr]
  * Annual cost: $0
]


---

# The Power of Parallel

To see the benefit of running code in parallel, let's write a .hi-blue[purposefully slow function:]

```{r}
slow_square <- function(x = 1){
  Sys.sleep(1/2) # wait half a second
  return(x^2)
}
```


---

# The Power of Parallel

How long does it take to run this function?<sup>5</sup>
  * Use `tic()` and `toc()` from .hi-slate[tictoc] to calculate elapsed time
  
```{r}
tic()
square_serial <- map(1:24, slow_square)
toc()
```
.footnote[<sup>5</sup> `sapply()` and `map()` take nearly the exact same time. There are also several [type-specific versions](https://www.rdocumentation.org/packages/purrr/versions/0.2.5/topics/map) of `map` in case you want output to be a logical, integer, double, or character, etc.]

--

The function runs in .hi-medgrn[serial]. so it takes approximately $1/2*24 = 12$ seconds
  * Using one core, runs for $x=1$, then when done moves on to $x=2, ..., 24$
  
---

# Parallelization

We can .hi-blue[speed this up]. Modern CPUs are made up of multiple .hi-pink[cores] (processing units) that can all be given tasks simultaneously, allowing us to run code in .hi-blue[parallel].

--

First, use `future::availableCores()` to determine how many cores you have:

```{r}
availableCores()
```


Your number of cores will likely differ
  * Most laptops have at least 4-8 cores these days.
  * Even recent Chromebooks have 6!
  
---

# furrr

.hi-slate[furrr] functions make it easy to .hi-blue[parallelize] in just a few steps.
  1. Set a "plan" for how the code will be run in parallel
    * Number of cores to use, how to execute tasks
  1. Use `future_` version of your preferred `map_` function
  1. Close parallel plan

--

First, we will .hi-medgrn[set the plan] and tell R how to execute the parallel session:

```{r}
# Calculate a "safe" number of cores (allow for background processes)
n_cores = availableCores() - 2

# Set the "plan"
plan(strategy = "multisession", # run in parallel in separate background R sessions
     workers = n_cores # use the desired number of cores
     )
```

---

# furrr

.hi-slate[furrr] functions make it easy to .hi-blue[parallelize] in just a few steps.
  1. Set a "plan" for how the code will be run in parallel
    * Number of cores to use, how to execute tasks
  1. Use `future_` version of your preferred `map_` function
  1. Close parallel plan


Next, let's repeat the previous analysis with `future_map()`.

```{r}
tic()
square_parallel <- future_map(1:24, slow_square)
toc()
```

---

# furrr

.hi-slate[furrr] functions make it easy to .hi-blue[parallelize] in just a few steps.
  1. Set a "plan" for how the code will be run in parallel
    * Number of cores to use, how to execute tasks
  1. Use `future_` version of your preferred `map_` function
  1. Close parallel plan


Now that we're done with our parallel session, reset things back to serial:

```{r}
plan("sequential")
```

---

# Benefits of Parallelization

Here we reduced execution time by ~ 1/3 due to some overhead of creating/assigning objects to the cores. However, the benefits of parallel increase substantially with
  * Larger objects
  * Greater number of repetitions (must be independent tasks)
  * More cores
  
--

For example, if we run our slow function over the integers 1 to 1,000:

| Approach | Time | Time Savings |
|----------|------|-------------------|
| Serial | 1,017.3 Seconds| 0%|
| Parallel, 5 cores | 204.33 Seconds | 80% |
| Parallel, 10 cores | 103.87 Seconds | 90% |
| Parallel, 20 cores | 51.49 Seconds| 95% |

```{r, include = F, eval = F}
# Running over a 1000 length vector in serial:
tic()
square_serial_long <- map(1:1000, slow_square)
toc()


# Run over 5 cores
plan(strategy = "multisession", # run in parallel in separate background R sessions
     workers = 5 # use the desired number of cores
     )

# Running over the same long vector in parallel:
tic()
square_parallel_long <- future_map(1:1000, slow_square)
toc()



# Run over 10 cores
plan(strategy = "multisession", # run in parallel in separate background R sessions
     workers = 10 # use the desired number of cores
     )

# Running over the same long vector in parallel:
tic()
square_parallel_long <- future_map(1:1000, slow_square)
toc()


# Run over 20 cores
plan(strategy = "multisession", # run in parallel in separate background R sessions
     workers = n_cores # use the desired number of cores
     )

# Running over the same long vector in parallel:
tic()
square_parallel_long <- future_map(1:1000, slow_square)
toc()

plan("sequential")
```

---
# Progress with *progressr*

For longer tasks, it can be helpful to see progress. We can do this by using the functions within .hi-slate[progressr].

First, let's add a .hi-purple[progress indicator] to our function.

```{r}
slow_square_prog <- function(x = 1){
  p() # add in progress indicator
  Sys.sleep(1/2) # wait half a second
  return(x^2)
}
```

---
# Progress with *progressr*

Next, write a .hi-medgrn[wrapper function] to our future map to add in the progress bar:

```{r}
par_slow_square <- function(x){
  p <- progressor(steps = length(x))
  future_map(x, slow_square_prog)
}
```

---
# Progress with *progressr*

Finally, wrap the function in `with_progress({})` to get a .hi-pink[visible progress bar].

```{r, eval = F}
with_progress({
  par_slow_square(1:24)
})
```
---
# Tweaking Progress Bar 

There are a lot of [different progress bar options](https://cran.r-project.org/web/packages/progressr/vignettes/progressr-intro.html), including
  * Change the shape used in the ASCII progress bar
```{r, eval = F}
pacman::p_load(cli)
handlers(handler_txtprogressbar(char = cli::col_red(cli::symbol$smiley)))

with_progress({
  par_slow_square(1:24)
})
```

---
# Tweaking Progress Bar 

There are a lot of [different progress bar options](https://cran.r-project.org/web/packages/progressr/vignettes/progressr-intro.html), including
  * Continuous color bar
```{r, eval = F}
handlers("cli")

with_progress({
  par_slow_square(1:24)
})
```



---
# Tweaking Progress Bar 

There are a lot of [different progress bar options](https://cran.r-project.org/web/packages/progressr/vignettes/progressr-intro.html), including
  * Audible beeps at start, intervals, and finish
```{r, eval = F}
pacman::p_load(beepr)
handlers("cli", "beepr")

with_progress({
  par_slow_square(1:24)
})
```



---
# Tweaking Progress Bar 

We can customize the sounds more fully with `handler_beepr()`:

```{r, eval = F}
sound_path <- paste0(getwd(), "/data/finish.wav")
handlers(list(
  "cli",
         handler_beepr(
           initiate = NA_integer_, # disable start sound
           update = NA_integer_, # disable progress sound
           finish = sound_path # set custom finish sound
            )
          )
        )
with_progress({
  par_slow_square(1:10)
})
```


---

# Table of Contents

1. [Prologue](#prologue)

2. [If/Else Statements](#ifelse)

3. [For Loops](#for)

4. [Functions](#fun)

5. [Indirection and Name Injection](#indirect)

6. [Vectorization](#vec)

7. [Parallelization](#parallel)





```{r gen_pdf, include = FALSE, cache = FALSE, eval = FALSE}
infile = list.files(pattern = 'Programming.html')
pagedown::chrome_print(input = infile, timeout = 2000)
```