---
title: "891/991 Final Presentations"
author: "James Sears <br> AFRE 891/991 SS 25 <br> Michigan State University"


      
output:
  xaringan::moon_reader:
    css: [msu-default.css, msu-metropolis.css, metropolis-fonts]
    lib_dir: libs
    self_contained: true
    nature:
      highlightStyle: github
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
      fig_caption: true

header-includes:
  -\usepackage{amsmath}
---

```{css, echo=FALSE}
# CSS for including pauses in printed PDF output (see bottom of lecture)
@media print {
  .has-continuation {
    display: block !important;
  }
}
.remark-code-line {
  font-size: 95%;
}
.small {
  font-size: 75%;
}
.scroll-output-full {
  height: 90%;
  overflow-y: scroll;
}
.scroll-output-75 {
  height: 75%;
  overflow-y: scroll;
}
```

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(knitr)
library(kableExtra)
library(fontawesome)
library(tidyverse)
knitr::opts_chunk$set(
	fig.align = "center",
	cache = FALSE,
	dpi = 300,
  warning = F,
  message = F,
	fig.height = 5,
	out.width = "80%"
)
```

# Table of Contents

1. [Final Presentation Structure](#structure)
1. [Final Presentation Schedule](#sched)
1. [Discussant Assignments](#discuss)


---
class: inverse, middle
name: structure

# Final Presentation Structure

---
# Final Presentation Structure

The final presentation is an .hi-purple[8 to 10 minute long, conference-style presentation]

* This is a .hi-medgrn[difficult length], as it is not enough time to convey everything
* Instead, focus on the .hi-pink[main story, overview of your approach, and the key results]

---
# Structure: Replication/Extension

.hi-medgrn[What to Include:]

  * Motivation
  * Statement of Replicated Paper
    * 991: statement of extensions
  * Overview of their Replication Package
  * Steps taken to replicate
    * Any translation from other languages
    * Comment on ease of replicability or if you ran into any issues/errors
  * 991: Discussion of extension results, implication for paper
  * Conclusion

---
# Structure: Acquisition

.hi-medgrn[What to Include:]

  * Motivation
  * Statement of Data Source
  * Overview of Necessary Approach
    * API: description of endpoints
    * Static/Dynamic: outline of workflow
  * Discussion of any challenges encountered
  * Summary of compiled dataset
  * Conclusion

---
# Structure: Analysis

.hi-medgrn[What to Include:]

  * Motivation
  * Research Question(s)
  * Summary of Data Sources
  * Empirical Strategy
  * Results
  * Conclusion/Next Steps

---
# Keep it Short

My recommendation: one slide per minute!

  * One slide each for motivation/conclusion/research statement
  * .hi-blue[Acquisition:] more than one for the workflow
  * .hi-medgrn[Analysis:] more than one for empirical strategy/results
  * .hi-purple[Replication/Extension:] more than one for steps taken/results
  
  
---
class: inverse, middle
name: sched

# Final Presentation Schedule

---
# Final Presentation Schedule

.pull-left[
.hi-medgrn.center[AFRE 991]

* Wednesday April 30
* 10am-12pm
* Our Scheduled Exam Time
* Ag Hall 75
]
--
.pull-left[
.hi-medgrn.center[AFRE 891: 2 Parts]

.pull-left[
.hi-purple.center[Group 1]
* Wednesday April 30
* 12:15pm-1:00pm
* Ag Hall 75
]

.pull-right[
.hi-purple.center[Group 2]
* Thursday May 1
* 10-10:30am
* Ag Hall 75
]

]

---
# 991 Presentation Schedule

.hi-medgrn.center[Wed April 30, Ag Hall 75]

```{r, echo = F}
set.seed(12345)
sched_991 <- data.frame(Presenter = c("Leslie Alex Joy",
"Dabit Bista",
"Mingda Li",
"Yixuan Li",
"Zhouxin Li",
"Lun Luo",
"Ashish Pokharel",
"Yuxin Shen"),
  Order = runif(n = 8)) %>%
  arrange(desc(Order)) %>%
  mutate(Order = row_number()) %>%
  mutate(Discussant = case_when(
    Order %in% seq(1,7,2) ~ lead(Presenter),
    Order %in% seq(2,8,2) ~ lag(Presenter))) %>%
  mutate(Time = c("10:00-10:12", "10:15-10:27", "10:30-10:42", "10:45-10:57",
                  "11:00-11:12", "11:15-11:27", "11:30-11:42", "11:45-11:57"
                  )) %>%    
  select(Order, Time, Presenter, Discussant) 



kable(sched_991, escape = F) %>% kable_styling(full_width = FALSE, bootstrap_options = c("striped", "bordered"))

```

---
# 891 Presentation Schedule


.pull-left[
.hi-medgrn.center[Group 1: Wed April 30, Ag Hall 75]

```{r, echo = F}
sched_891_1 <- data.frame(Presenter = c("Caitlin Nelson",
"Michael Hall",
"Sam Rickman"),
  Order = runif(n = 3)) %>%
  arrange(desc(Order)) %>%
  mutate(Order = row_number()) %>%
  mutate(Discussant = case_when(
    Order %in%(1:2) ~ lead(Presenter),
    TRUE ~ lag(Presenter,2))) %>%
  mutate(Time = c("12:15-12:27", "12:30-12:42", "12:45-12:57")) %>%    
  select(Order, Time, Presenter, Discussant) 



kable(sched_891_1, escape = F) %>% kable_styling(full_width = FALSE, bootstrap_options = c("striped", "bordered"))

```

]

.pull-right[
.hi-blue.center[Group 2: Thur May 1, Ag Hall 75]

```{r, echo = F}
sched_891_2 <- data.frame(Presenter = c("Kelly Hudson",
"Ye Rim Lee"),
  Order = runif(n = 2)) %>%
  arrange(desc(Order)) %>%
  mutate(Order = row_number()) %>%
  mutate(Discussant = case_when(
    Order == 1 ~ lead(Presenter),
    TRUE ~ lag(Presenter))) %>%
  mutate(Time = c("10:00-10:12", "10:15-10:27")) %>%    
  select(Order, Time, Presenter, Discussant) 



kable(sched_891_2, escape = F) %>% kable_styling(full_width = FALSE, bootstrap_options = c("striped", "bordered"))

```

]

---
# Scheduled Times

I put presenters + discussants in .hi-medgrn[back-to-back half hour windows].


If you have a conflict with the above scheduled time such that you .hi-medgrn[absolutely can't make that half hour window], then .hi-blue[find another pair to switch with] and then .hi-blue[confirm the switch with me].


---
# Replication Packages

Remember that replication packages must be submitted to .hi-medgrn[myself and your discussant] by .hi-slate[end of day Sunday, April 20].

  * Late submissions on replication packages will .hi-slate[not receive points] (7% of final grade)
  
You can either
  1. Create a GitHub Repository and share it that way, or
  2. Email a (link to a) zip folder containing all replication materials


```{r gen_pdf, include = FALSE, cache = FALSE, eval = FALSE}
infile = list.files(pattern = '991.html')
pagedown::chrome_print(input = infile, timeout = 200)
```