---
title: "Lecture 9: Synthetic Control Methods, Part 1"
author: "James Sears* <br> AFRE 891/991 SS 25 <br> Michigan State University"
date: ".small[<br> *Parts of these slides are adapted from [“Econometrics III”](https://github.com/edrubin/EC607S23) by Ed Rubin.]"

      
output:
  xaringan::moon_reader:
    css: [msu-default.css, msu-metropolis.css, metropolis-fonts]
    lib_dir: libs
    latex_engine: lualatex
    self_contained: true
    nature:
      highlightStyle: github
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
      fig_caption: true

header-includes:
  -\usepackage{amsmath}
---

```{css, echo=FALSE}
# CSS for including pauses in printed PDF output (see bottom of lecture)
@media print {
  .has-continuation {
    display: block !important;
  }
}
.remark-code-line {
  font-size: 95%;
}
.small {
  font-size: 75%;
}
.scroll-output-full {
  height: 90%;
  overflow-y: scroll;
}
.scroll-output-75 {
  height: 75%;
  overflow-y: scroll;
}
```

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(knitr)
library(fontawesome)

# define colors
red_pink <- "#e64173"
turquoise <- "#20B2AA"
orange <- "#FFA500"
red <- "#fb6107"
blue <- "#3b3b9a"
green <- "#8bb174"
grey_light <- "grey70"
grey_mid <- "grey50"
grey_dark <- "grey20"
purple <- "#6A5ACD"
slate <- "#314f4f"

knitr::opts_chunk$set(
	fig.align = "center",
	cache = FALSE,
	dpi = 300,
  warning = F,
  message = F,
	fig.height = 5,
	out.width = "80%"
)
```

# Table of Contents


1. [Prologue](#prologue)

3. [Matching](#match)

4. [Canonical Synthetic Control](#canon)

---
class: inverse, middle
name: prologue

# Prologue

---
# Prologue

This lecture is focusing on .hi-medgrn[Synthetic Control Methods], which will let us solve several of the issues that can affect methods we discussed last lecture.

--

.pull-left[
.center.hi-purple[Part 1]
  * The Fundamental Problem of Causal Inference
  * Matching
  * Canonical Synthetic Control
]

--

.pull-right[
.center.hi-pink[Part 2]
  * Synthetic Diff-in-Diff
    * Uniform Adoption
    * Staggered Adoption
  * Partially Pooled Synthetic Control
]

---
# Prologue

Packages we'll use today:


```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(fixest,  tidyverse)
```

--

As well, let's load the event study data from [Sears et al. (2023)](https://doi.org/10.1086/721705) we finished last lecture with:

```{r}
sah <- readRDS("data/sah_es.rds")
```

---
# Causal Inference

Let's chat about the .hi-blue[fundamental problem of causal inference] for a moment.

Consider unit $i$'s .hi-medgrn[potential outcomes:]
--

  * $\color{#3b3b9a}{\mathbf{Y}_{1i}}$: the outcome for unit $i$ .hi-blue[under the treatment]
    
    * Treatment assignment $D_i = 1$
--
  
  * $\color{#fb6107}{\mathbf{Y}_{0i}}$: the outcome for unit $i$ .hi-red[absent the treatment]
    
    * Treatment assignment $D_i = 0$

---
# Causal Inference

* $\color{#3b3b9a}{\mathbf{Y}_{1i}}$: the outcome for unit $i$ .hi-blue[under the treatment]
* $\color{#fb6107}{\mathbf{Y}_{0i}}$: the outcome for unit $i$ .hi-red[absent the treatment]

<br>

We want to know 

$$\tau_i = \text{Y}_{1i} - \text{Y}_{0i}$$

--
<br>

.hi-blue[Fundamental problem:] we cannot simultaneously observe *both* $\text{Y}_{1i}$ *and* $\text{Y}_{0i}$.

--

Most (all?) empirical strategies boil to .hi-red[estimating] $\color{#fb6107}{\mathbf{Y}_{0i}}$ for treated individuals — the .hi-green[unobservable counterfactual] for the treatment group.

---
# Causal Inference

Last lecture gave an overview of regression methods that make different assumptions about that .hi-green[unobservable counterfactual].

--

.hi-blue[1\. RCT + Random Assignment]
  * The average in the control group is what the average in the treatment group would have been absent the treatment
  * On average, no systematic difference in (un)observables between treatment + control
  * Regress outcome on treatment dummy and you're good to go
    * Maybe add some control variables to improve precision of estimator
--



.hi-medgrn[2\. Difference-in-Differences + Event Study]

  * The .hi-medgrn[change over time] in the control group is what the change in the treatment group would have been absent the treatment

---
# Causal Inference


All of these estimates are identified under a variation of the .hi-purple[Conditional Independence Assumption (CIA)].super[.pink[†]] 



$$
\begin{align}
  \def\ci{\perp\mkern-10mu\perp}
  \left\{ \color{#6A5ACD}{\text{Y}_{0i}},\,\color{#6A5ACD}{\text{Y}_{1i}} \right\} \ci  \color{#e64173}{\text{D}_{i}} ~|~ \text{X}_{i}
\end{align}
$$
--

Conditional on $\text{X}_{i}$.super[1], potential outcomes $\left( \color{#6A5ACD}{\text{Y}_{0i}},\, \color{#6A5ACD}{\text{Y}_{1i}} \right)$ are independent of treatment status $\left( \color{#e64173}{\text{D}_{i}} \right)$.


.footnote[.pink[†] AKA "selection on observables". <br> 1. Or $_{it}$ if we're in a panel setting]

---
# Causal Inference

But there are times when .hi-pink[CIA fails].super[2].

.footnote[2\. See the [Bay of Pigs Invasion](https://www.jfklibrary.org/learn/about-jfk/jfk-in-history/the-bay-of-pigs)]

In the case of Diff-in-Diff and Event Study, this is often due to a failure of .hi-purple[parallel trends].

--

For example, recall the event study for mobility responses to stay-at-home mandates from last lecture:

```{r, echo = F, out.width = "50%"}
es_reg <- feols(cadt ~ i(event_time_bin, sah_state, ref = -1) | 
                  state + date, data = sah)
iplot(es_reg,
      ylab = "Change in Average Distance Traveled (%)",
      xlab = "Days Since Stay-at-Home in Effect",
      main = "Event Study, Average Distance Traveled"
)
```

---
# Causal Inference

```{r, echo = F, out.width = "60%"}
es_reg <- feols(cadt ~ i(event_time_bin, sah_state, ref = -1) | 
                  state + date, data = sah)
iplot(es_reg,
      ylab = "Change in Average Distance Traveled (%)",
      xlab = "Days Since Stay-at-Home in Effect",
      main = "Event Study, Average Distance Traveled"
)
```

In this case, the states that never adopted stay-at-home mandates might not be *valid counterfactuals* for the states that did adopt.

--

However, there might be a way to *construct* a valid counterfactual from the set of control units...


---
# Causal Inference


```{r, echo = F, out.width = "60%"}
es_reg <- feols(cadt ~ i(event_time_bin, sah_state, ref = -1) | 
                  state + date, data = sah)
iplot(es_reg,
      ylab = "Change in Average Distance Traveled (%)",
      xlab = "Days Since Stay-at-Home in Effect",
      main = "Event Study, Average Distance Traveled"
)
```

<br>
... but before we get into that, let's chat briefly about .hi-purple[matching estimators].

---
class: inverse, middle
name: match

# Matching

---

# Matching

.hi-purple[Matching Estimators] provide an alternate way of coming up with the unobservable counterfactual for the treatment group.

--

.hi-medgrn[The gist:]

  * Match untreated observations to treated observations similar on $\text{X}_{i}$
    * _i.e._ calculate a $\widehat{\text{Y}_{0i}}$ for each $\text{Y}_{1i}$, based upon "matched" untreated individuals with (nearly) identical values of $X_i$
--
  * If CIA holds, then we can just calculate a bunch of treatment effects conditional on $\text{X}_{i}$
    * _i.e._ 
    
$$
\begin{align}
  \tau(x) = \mathop{E}\left[ \text{Y}_{1i} - \text{Y}_{0i} \mid \text{X}_{i} = x \right]
\end{align}
$$ 

---

# Matching (Formally)


We want to construct a counterfactual for each individual with $\text{D}_{i}=1$.

--

.hi-medgrn.note[CIA:] The counterfactual for $i$ should .hi-medgrn[only use individuals that match] $\color{#008208}{\text{X}_{i}}$.

--

Let there be $N_T$ treated individuals and $N_C$ control individuals. We want
--

* $N_T$ sets of weights
* $N_C$ weights in each set


$$w_i(j)\, \left( i = 1,\,\ldots,\, N_T;\, j=1,\,\ldots,\, N_C \right)$$

--

Assume $\sum_j w_i(j) = 1$. Our estimate for the counterfactual of treated $i$ is
$$
\begin{align}
  \widehat{\text{Y}_{0i}} = \sum_{j\in \left( D=0 \right)} w_i(j) \text{Y}_{j}
\end{align}
$$

---
# Weight for it

So all we need is those weights and we're done.

.hi-medgrn[Q:] Where does one find these handy weights?

--

.hi-blue[A:] You've got options, but you need to choose carefully/responsibly.

*E.g.* if $w_i(j) = \frac{1}{N_C}$ for all $(i,j)$, then we're back to a .hi-pink[difference in means].
<br> This weighting doesn't abide by our conditional independence assumption.

--

.hi-medgrn[The plan:] choose weights $w_i(j)$ that indicate .hi-slate[*how close*] $\text{X}_{j}$ is to $\text{X}_{i}$.

---
# Weight for it

Some common choice of weights:

  * .hi-medgrn[Nearest neighbor:]
$$
\begin{align}
  \text{d}_{i,j} = \left( \text{X}_{i} - \text{X}_{j} \right)'\left(\text{X}_{i} - \text{X}_{j}\right)
\end{align}
$$
  * .hi-blue[Kernel Matching] for .hi-slate[bandwidth] $h$ and .hi-slate[kernel function] $K(\cdot)$:
$$
\begin{align}
  w_i(j) = \dfrac{K\!\!\left( \dfrac{\text{X}_{j} - \text{X}_{i}}{h} \right)}{\sum\limits_{j\in(D=0)} K\!\!\left(\dfrac{\text{X}_{j} - \text{X}_{i}}{h} \right)}
\end{align}
$$

---
# Kernels

For example, the .hi-blue[Epanechnikov kernel] is defined as
$$
\begin{align}
  K(z) = \dfrac{3}{4} \left( 1 - z^2 \right) \times \mathbb{I}\!\left( |z| < 1 \right)
\end{align}
$$

```{r, epanechnikov, echo = F, out.width = "70%"}
ggplot(
  data = data.frame(x = c(-2.5, 2.5)),
  aes(x = x)
) +
geom_hline(
  yintercept = 0,
  color = "grey70"
) +
stat_function(
  fun = function(x) 3/4 * (abs(x) <= 1) * (1 - x^2),
  color = orange,
  size = 2.5
) +
ylim(0, 1) +
xlab("z") +
ylab("K(z)") +
theme_minimal(base_size = 20, base_family = "Fira Sans Book") +
theme(
  axis.title = element_text(family = "STIXGeneral", face = "italic", size = 22),
  panel.grid.major = element_line(size = 0.5, color = "grey95"),
  panel.grid.minor = element_line(size = 0.5, color = "grey95")
)
```

---
# Kernels

And the .hi-red[triangular kernel] can be expressed as  $K(z) = \left( 1 - |z| \right) \times \mathbb{I}\!\left( |z| < 1 \right)$

```{r, triangle, echo = F}
ggplot(
  data = data.frame(x = c(-2.5, 2.5)),
  aes(x = x)
) +
geom_hline(
  yintercept = 0,
  color = "grey70"
) +
stat_function(
  fun = function(x) (abs(x) <= 1) * (1 - abs(x)),
  color = orange,
  size = 2.5
) +
ylim(0, 1) +
xlab("z") +
ylab("K(z)") +
theme_minimal(base_size = 20, base_family = "Fira Sans Book") +
theme(
  axis.title = element_text(family = "STIXGeneral", face = "italic", size = 22),
  panel.grid.major = element_line(size = 0.5, color = "grey95"),
  panel.grid.minor = element_line(size = 0.5, color = "grey95")
)
```
---
# Kernels

And the .hi-purple[uniform kernel] with $K(z) = \frac{1}{2} \times \mathbb{I}\!\left( |z| < 1 \right)$

```{r, uniform, echo = F}
ggplot(
  data = data.frame(x = c(-2.5, 2.5)),
  aes(x = x)
) +
geom_hline(
  yintercept = 0,
  color = "grey70"
) +
stat_function(
  fun = function(x) (abs(x) <= 1) * 1/2,
  n = 1e3,
  color = orange,
  size = 2.5
) +
ylim(0, 1) +
xlab("z") +
ylab("K(z)") +
theme_minimal(base_size = 20, base_family = "Fira Sans Book") +
theme(
  axis.title = element_text(family = "STIXGeneral", face = "italic", size = 22),
  panel.grid.major = element_line(size = 0.5, color = "grey95"),
  panel.grid.minor = element_line(size = 0.5, color = "grey95")
)
```
---
# Kernels

Or the .hi-pink[Gaussian kernel] $K(z) = \left( 2\pi \right)^{-1/2}  \exp\left(-z^2/2 \right)$

```{r, gaussian, echo = F}
ggplot(
  data = data.frame(x = c(-2.5, 2.5)),
  aes(x = x)
) +
geom_hline(
  yintercept = 0,
  color = "grey70"
) +
stat_function(
  fun = function(x) (2 * pi)^(-1/2) * exp(-x^2/2),
  color = orange,
  size = 2.5
) +
ylim(0, 1) +
xlab("z") +
ylab("K(z)") +
theme_minimal(base_size = 20, base_family = "Fira Sans Book") +
theme(
  axis.title = element_text(family = "STIXGeneral", face = "italic", size = 22),
  panel.grid.major = element_line(size = 0.5, color = "grey95"),
  panel.grid.minor = element_line(size = 0.5, color = "grey95")
)
```
---
# Aside on Kernels

Kernel functions are good for more than just matching.

You will most commonly see/use them smoothing out densities—providing a smooth, moving-window average.

--

_E.g._ .mono[R]'s (`ggplot2`'s) smooth, density-plotting function `geom_density()`.

`geom_density()` defaults to `kernel = "gaussian"`, but you can specify many other kernel functions (including `"epanechnikov"`).

--

You can also change the `bandwidth` argument. The default is a bandwidth-choosing function called `bw.nrd0()`.

---
# Matching

More recent techniques approach matching in one of two ways

1\. As a .hi-medgrn[subset selection process]
  * Use algorithms to produce (weighted) subsets where treatment is unassociated with (un)observed covariates
    * Treated units paired with control units; unpaired units are dropped (stratification)
  * Reduce reliance on correct model specification in subsequent regressions (Ho et al. 2007)
  * A non-parametric pre-processing approach 

---
# Matching

More recent techniques approach matching in one of two ways

2\. As a .hi-purple[counterfactual imputation method]
  * Pair units
  * Impute the missing potential outcome for each unit using the observed outcomes of paired units
  * Well-understood theoretical properties (Abadie and Imbens 2006, 2016)
  
---
# Matching

If taking a .hi-medgrn[subset selection] approach, there are many options:

* .hi-slate[Nearest Neighbor (Greedy)]: pair to control unit(s) closest on a distance measure
--

* .hi-slate[Optimal Pair]: pair to control unit(s) by optimizing an overall criterion
--

* .hi-slate[Optimal Full]: Assign each treated + control units to a subclass
  * each subclass: 1 T and $\geq$1 C, or $\geq$1 T and 1 C
  * Minimizes the within-subclass distances
  * Weights chosen by subclass membership
  * .hi-slate[Generalized Full]: fast clustering algorithm to improve comp. time
--

* .hi-slate[Genetic]: find scaling factors for each covariate, then do NN

---
# Matching

If taking a .hi-medgrn[subset selection] approach, there are many options:


* .hi-slate[Exact]: create subclasses using unique covariate values
  * Exactly balances covariate distributions
  * .hi-slate[Coarsened Exact]: exact but on binned covariate values
--

* .hi-slate[Cardinality/Profile]: "all or nothing" weighting method
    * Finds the largest sample that satisfies provided balance constraints
    * Solves a mixed integer program ([Zubizerreta, Paredes, and Rosenbaum 2014](http://jrzubizarreta.com/cardinality.pdf))
    * Cardinality: balance on mean differences in matched sample, T + C samples same size
    * Profile: balance on mean difference between T and target distribution
    * Could then also create pairs to improve precision ([Zubizerreta, Paredes, and Rosenbaum 2014](http://jrzubizarreta.com/cardinality.pdf))

---
# Matching

Woah, that's a lot of options. How do we choose?

--

[Ho et al. (2007)](https://doi.org/10.1093/pan/mpl013) recommend the following:


--

* Try a matching method
  * For ATE: optimal, generalized full, subclassification, or profile 
  * For ATT/ATC: any
* Assess balance
  * Distributional balance, not just mean balance
* If it yields poor balance or a too-low sample size, try another method
* Attempt to come as close as possible to perfect balance

---
class: inverse, middle
name: canon

# Canonical Synthetic Control

---
# Canonical Synthetic Control

The .hi-medgrn[canonical synthetic control method] feels a lot like if .hi-blue[event study] and a .hi-green[matching estimator] got together and had a kid.

--

  * Originated in [Abadie and Gardeazabal (2003) AER](https://www.aeaweb.org/articles?id=10.1257/000282803321455188), refined and extended in [Abadie, Diamond, and Hainmueller (2010) JASA](https://www.tandfonline.com/doi/abs/10.1198/jasa.2009.ap08746)
  
  * Developed for .hi-medgrn[comparative case studies:] one aggregate unit exposed to treatment/intervention

--

.hi-purple[The gist:] compare post-treatment outcome evolution in treated group to a .hi-slate[synthetic control] unit constructed to match on 
  
  * Pre-trends in the outcome variable $\text{Y}_i$,
  * and optionally: Covariates $\text{X}_{i}$ or $\text{X}_{it}$

---
# Canonical Synthetic Control

Synthetic control overcomes one of the key issues of studying aggregate units: .hi-medgrn[few, poor-counterfactual controls]

--

  * Policy interventions often happen at an aggregate level (i.e. state, country)
  * Aggregate/macro data are often easy to obtain

However,

  * Finding a valid counterfactual with coarse, aggregate units can be difficult
  * Control group selection is ad hoc, leading to *researcher degrees of freedom*
  

---
# Canonical Synthetic Control

.hi-medgrn[Formally:]
  * Suppose you have data for $J+1$ units
    * Treated unit: $j=1$
    * "Donor Pool": all $j=2,...,J+1$ units
  * Data span $T$ periods, with $T_0$ periods prior to treatment

--

For each unit, we observe
1. The outcome of interest $Y_{jt}$
--

1. A set of $k$ predictors of the outcome, $X_{1j},...,X_{kj}$
    * May include pre-intervention values of $Y_{jt}$
    * Must be .hi-blue[unaffected by the intervention]

---
# Canonical Synthetic Control

.hi-medgrn[Formally:]

For each unit $j$, let $Y_{jt}^N$ be the potential response .hi-pink[without intervention], with $Y_{1t}^I$ the potential response .hi-blue[under intervention] for the treated unit
    
  * For unit "one" with $t > T_0$, we have $Y_{1t} = Y_{1t}^I$
      
Under this setup, the effect of the policy in period $t$ is given by 

$$\tau_{1t} = Y_{1t}^I - Y_{1t}^N$$

--

.hi-purple[Policy Evaluation Challenge:] how to estimate $Y_{1t}^N$, the unobserved counterfactual?

---
# Canonical Synthetic Control

.hi-blue[A:] Construct a .hi-medgrn[synthetic control] as a weighted average of units in the donor pool.

--

Let $W = (\omega_2,...,\omega_{J+1})'$ be a $J \times 1$ vector of weights.

--

For a given $W$, the synthetic control counterfactual is

  $$\hat{Y}_{1t}^N = \sum\limits_{j=1}^{J+1} \omega_jY_{jt}$$
  
  and 

$$\hat\tau_{1t} = Y_{1t}^I - \hat Y_{1t}^N$$
---
# Choosing Weights

Weights are designed to .hi-blue[avoid extrapolation]
  
* $\omega_j \geq 0 ~~\forall~ j$

* $\sum\limits_{j=1}^{J+1} \omega_j = 1$

* Ensures synthetic control is located within the convex hull of donor units (based purely on observed data)
  
--

We will choose the $\omega_j$ so that the synthetic control   .hi-medgrn[matches pre-intervention values for the treated unit of predictors for the outcome variable].

---
# Identification Under SCM

.hi-medgrn[matches pre-intervention values for the treated unit of predictors for the outcome variable].

The whole idea behind SCM and its nice statistical properties relies on our ability to .hi-medgrn[approximate relevant unobservables].

--

[Arkhangelsky and Hirshberg (2023)](https://arxiv.org/pdf/2311.13575) formalize that identification of canonical SCM *and* the extensions we'll talk through next time holds under .hi-blue[two main conditions:]

* .hi-purple[Sequential Exogeneity:] conditional on observed pre-treatment information and permanent unobserved characteristics, treatment is independent of future counterfactual outcomes.
  * .hi-purple[Best case:] conditional mean is a linear function of past outcomes (e.g. random walk)


---
# Identification Under SCM

.hi-medgrn[matches pre-intervention values for the treated unit of predictors for the outcome variable].

The whole idea behind SCM and its nice statistical properties relies on our ability to .hi-medgrn[approximate relevant unobservables].



[Arkhangelsky and Hirshberg (2023)](https://arxiv.org/pdf/2311.13575) formalize that identification of canonical SCM *and* the extensions we'll talk through next time holds under .hi-blue[two main conditions:]

* .hi-pink[Retrievability:] when the number of pre-treatment periods is sufficiently large, relevant unobservables can be precisely recovered.

---
# Identification Under SCM

If  .hi-purple[Sequential Exogeneity] and  .hi-pink[Retrievability] hold, then
  * SCM estimators are consistent and asymptotically normal, even when treatment assignment is a deterministic function of the outcome (permanent and/or time-varying)
    * DiD: *not* consistent here

---
# Identification Under SCM

Situations where .hi-red[SCM Breaks Down]:

--

* Heterogeneity in the persistence of time-varying shocks
  * $\Rightarrow$ pre-treatment information insufficient to construct weights
  * $\Rightarrow$ potentially overcome by including unit-level measures of persistence

--
* Violation of strict exogeneity
  * $\Rightarrow$ adoption decisions are correlated with time-varying component of outcome  (conditional on permanent unobserved heterogeneity)

---
# Choosing Weights

Formally, choose weights $W^*$ that minimize


$$||X_1 - X_0W || = (\sum\limits_{h=1}^k \nu_h(X_{h1} - \omega_2X_{h2} - ... - \omega_{J+1}X_{hJ+1})^2)^{1/2}$$

* Positive constants $\nu_1...\nu_k$ reflect the .hi-pink[relative importance] put on predictors $1,...k$

--

  * [Abadie, Diamond, and Hainmueller (2010)](https://www.tandfonline.com/doi/abs/10.1198/jasa.2009.ap08746): select $\nu_1...\nu_k$ to minimize .hi-red[mean square prediction error (MSPE)] for some set of pre-intervention periods



---
# Choosing Weights

Formally, choose weights $W^*$ that minimize


$$||X_1 - X_0W || = (\sum\limits_{h=1}^k \nu_h(X_{h1} - \omega_2X_{h2} - ... - \omega_{J+1}X_{hJ+1})^2)^{1/2}$$

* Positive constants $\nu_1...\nu_k$ reflect the .hi-pink[relative importance] put on predictors $1,...k$



  * [Abadie, Diamond, and Hainmueller (2015)](https://onlinelibrary.wiley.com/doi/full/10.1111/ajps.12116): select $\nu_1...\nu_k$ via .hi-bluep[out-of-sample validation]
    1. Divide pre-intervention period into  .hi-medgrn[training] and .hi-purple[validation] periods
    1. Select a value of $V^* = \nu^*_1...\nu^*_k$ that yields a small MSPE in the validation period
    1. Use resulting $V^*$ to calculate optimal weights in the validation period 
    
    
---
# German Reunification
To see how this process works, let's see an example: .hi-slate[Impact of 1990 German Reunification on GDP].

Let's load in some state-by-year data on GDP and other economic conditions:

```{r}
deu <- haven::read_dta("data/repgermany.dta") %>%
    mutate_at(vars(index, year, gdp, infrate, trade, schooling,
                   invest60, invest70, invest80,
                   industry),
               as.numeric) %>%
    mutate_at(vars(country), as.character)
```

---
# German Reunification

Let's load in some state-by-year data on GDP and other economic conditions:

```{r}
head(deu)
```

---
# German Reunification

How did West Germany GDP compare to OECD countries .hi-blue[prior] to reunification in 1990?

  
```{r, echo = F, out.width = "65%"}
filter(deu, year <= 1990) %>%
  mutate(wg = ifelse(country == "West Germany", "West Germany", "OECD")) %>%
  group_by(wg, year) %>%
  summarise(gdp = mean(gdp, na.rm = T)) %>%
  ggplot() +
  geom_line(aes(x = year, y = gdp, linetype = wg)) +
  labs(x = NULL, y = "Per Capita GDP (2002 USD)", linetype = NULL) +
  theme_minimal() +
  lims(y = c(0, 21000))  
```

--
  
  * _Spoiler:_ that gap looks to be .hi-medgrn[growing]



---
# German Reunification

What if we construct a "synthetic" West Germany to match on pre-unification predictors of economic growth?

  * Pre-treatment outcome
    * GDP (average for 1981-1990)

--
  
  * Covariates
    * Trade openness: Exports + Imports as % of GDP (average for 1981-1990)
    * Inflation Rate (average for 1981-1990)
    * Industry share of value-added (average 1971-1980)
    * Schooling: % of secondary school attained in the age 25+ population (5-year averages, 1970 and 1975)
    * Investment rate: ratio of real domestic investment (private + public) to real GDP  (5-year average 1970)



---
# German Reunification

Let's use the .hi-slate[tidysynth] package to do this in a *tidy* workflow

--

.hi-purple[one hiccup:] .hi-slate[tidysynth] was recently .hi-purple[removed from CRAN]<sup>1</sup>

```{r, error = T}
install.packages("tidysynth")
```


.footnote[<sup>1</sup> CRAN has automatic checks in place that sometimes remove packages because of seemingly arbitrary reasons. .hi-slate[tidysynth] was removed because it failed some automated HTML manual issues, not because of any issue with the methods the package uses.]

---
# Installing from Github

We can get around this pretty easily by remembering the  .hi-slate[remotes] package we saw in lecture 1
  * Allows installation of packages directly from
    * GitHub, GitLab, BitBucket
    * Local files
    * Specific package versions on CRAN

```{r}
remotes::install_github("edunford/tidysynth")
pacman::p_load(tidysynth)
```


---
# German Reunification

Let's use the .hi-slate[tidysynth] package to do this in a *tidy* workflow




First, let's set up the synthetic control object with `synthetic_control()`
<br>

```{r}
synth_wg <- deu %>%
    synthetic_control(
      outcome = gdp, # outcome
      unit = country, # unit identifier
      time = year, # time identifier
      i_unit = "West Germany", # treated unit
      i_time = 1990, # treatment year
      generate_placebos = T # whether to generate placebos for inference
    )
```


---
# German Reunification

Next, add the predictors with `generate_predictor()`

  * Choose a time period to use for matching (`time_window`)
    * Default: full pre-intervention period
  * Choose the variables to use 
  * Choose the summary method (e.g. `sum()`, `mean()`, `min()`, `sum(is.na())`)
  * Choose the variable names for the new predictors

---
# German Reunification

Adding the predictors with `generate_predictor()`:

  * First: 1981-1990 averages for GDP, Trade Openness, and Inflation Rates
```{r, eval = F}
synth_wg <- synth_wg %>%
  generate_predictor(time_window = 1981:1990, #<<
                     gdp_81_90 = mean(gdp, na.rm = T), #<<
                     trade81_90 = mean(trade, na.rm = T), #<<
                     infrate81_90 = mean(infrate, na.rm = T) #<<
  ) %>%
  generate_predictor(time_window = 1971:1980,
                     industry_71_80 = mean(industry, na.rm = T)) %>%
  generate_predictor(time_window =c(1970, 1975),
                     schooling_70_75 = mean(schooling, na.rm = T)) %>%
  generate_predictor(time_window = 1980,
                     invest_70 = invest70)
```


---
# German Reunification

Adding the predictors with `generate_predictor()`:

  * Second: 1971-1980 average for Industry shares
```{r, eval = F}
synth_wg <- synth_wg %>%
  generate_predictor(time_window = 1981:1990, 
                     gdp_81_90 = mean(gdp, na.rm = T),
                     trade81_90 = mean(trade, na.rm = T), 
                     infrate81_90 = mean(infrate, na.rm = T)
  ) %>%
  generate_predictor(time_window = 1971:1980, #<<
                     industry_71_80 = mean(industry, na.rm = T)) %>% #<<
  generate_predictor(time_window =c(1970, 1975),
                     schooling_70_75 = mean(schooling, na.rm = T)) %>%
  generate_predictor(time_window = 1980,
                     invest_80 = invest80)
```


---
# German Reunification

Adding the predictors with `generate_predictor()`:

  * Mean of 1970-1975 and 1975-1980 secondary school attainment
  
```{r}
synth_wg <- synth_wg %>%
  generate_predictor(time_window = 1981:1990, 
                     gdp_81_90 = mean(gdp, na.rm = T),
                     trade81_90 = mean(trade, na.rm = T), 
                     infrate81_90 = mean(infrate, na.rm = T)
  ) %>%
  generate_predictor(time_window = 1971:1980, 
                     industry_71_80 = mean(industry, na.rm = T)) %>% 
  generate_predictor(time_window =c(1970, 1975), #<<
                     schooling_70_75 = mean(schooling, na.rm = T)) %>% #<<
  generate_predictor(time_window = 1980,
                     invest_80 = invest80)
```

---
# German Reunification

Adding the predictors with `generate_predictor()`:

  * 1980-1985 ratio of (real) domestic investment to GDP
  
```{r, eval = T}
synth_wg <- synth_wg %>%
  generate_predictor(time_window = 1981:1990, 
                     gdp_81_90 = mean(gdp, na.rm = T),
                     trade81_90 = mean(trade, na.rm = T), 
                     infrate81_90 = mean(infrate, na.rm = T)
  ) %>%
  generate_predictor(time_window = 1971:1980, 
                     industry_71_80 = mean(industry, na.rm = T)) %>% 
  generate_predictor(time_window =c(1970, 1975),
                     schooling_70_75 = mean(schooling, na.rm = T)) %>% 
  generate_predictor(time_window = 1980, #<<
                     invest_80 = invest80) #<<
```


---
# German Reunification

Next, generate weights with `generate_weights()`

```{r}
wts <- synth_wg %>%
  generate_weights(optimization_window = 1981:1990)

# get variable weights
wt_vec <- wts[[7]][[1]] %>%
  select(weight) %>% as.vector() %>% unlist()
```


---
# German Reunification

Finally, estimate the synthetic control and plot it
```{r}
synth_control <- generate_control(wts)

plot_trends(synth_control)
```

---
# German Reunification

Alternatively, extract the synthetic control + treated unit values for manual plotting:

```{r}
grab_synthetic_control(synth_control) %>% head()
```

---
# German Reunification

```{r, out.width = "60%"}
grab_synthetic_control(synth_control) %>% 
  pivot_longer(cols = ends_with("y"), names_to = "var") %>%
  ggplot(aes(x = time_unit)) +
  geom_line(aes(y = value, linetype = var)) +
    geom_vline(aes(xintercept = 1990), linetype = "dashed") +
  theme_minimal()
```

---
# German Reunification

Comparing to the raw mean of OECD countries:

```{r, echo = F, out.width = "80%"}
  mutate(deu, wg = ifelse(country == "West Germany", "West Germany", "OECD")) %>%
  group_by(wg, year) %>%
  summarise(gdp = mean(gdp, na.rm = T)) %>%
  ggplot() +
  geom_line(aes(x = year, y = gdp, linetype = wg)) +
  geom_vline(aes(xintercept = 1990), linetype = "dashed") +
  labs(x = NULL, y = "Per Capita GDP (2002 USD)", linetype = NULL) +
  theme_minimal() +
  lims(y = c(0, 35000))  
```

---
# German Reunification

Alernatively we can plot the .hi-blue[difference] between West Germany and its synthetic control:

```{r, out.width = "65%"}
plot_differences(synth_control)
```


---
# German Reunification


Looking at both the control unit and variable weights with `plot_weights()`:

```{r, out.width = "70%"}
synth_control %>%
  plot_weights() #<<
```

---
# German Reunification

Checking balance of real West Germany vs. Synthetic West Germany vs. Mean of OECD Countries with `grab_balance_table()`: 
```{r}
synth_control %>%
  grab_balance_table() #<<
```

---
# German Reunification

For inference, we repeat the same process as before with every unit in the donor pool.

```{r, out.width = "75%"}
synth_control %>%
  plot_placebos(prune = FALSE)
```

---
# German Reunification

By default, `plot_placebos()` hides the placebo controls with large MSPEs (here we only get 3)

```{r, out.width = "75%"}
synth_control %>%
  plot_placebos()
```
---

# Inference

Finally, looking at inference:

```{r}
wg_inf <- synth_control %>%
  grab_significance() #<<
wg_inf
```
---

# Inference

```{r}
colnames(wg_inf)
```

Inference with synthetic controls is based on the .hi-blue[difference between pre and post-intervention MSPE] values.

.hi-medgrn[Idea:] if the synthetic control fits the observed data well (low pre-intervention MSPE), and diverges in the post-period (high post-period MSPE), then the intervention had a meaningful effect.

  * If the intervention had *no* effect, the pre and post-period MSPE should be similar, with a ratio around 1
  * If placebos fit the data as well as the treated unit, we can't reject the null of no treatment effect
---

# Inference

.hi-purple[Fisher's exact P-value] is generated by first ranking ratios then dividing the rank of the case over the total

```{r}
unique_countries <- unique(deu$country) %>% length()

# Fisher's P calculated as rank/total, so for West Germany (rank 1):
which(wg_inf$unit_name == "West Germany")/unique_countries
```

Z-score is then the standardized RMSE ratios for all cases.
  
  * Captures degree to which a particular case's RMSE ratio deviates from the .hi-pink[placebo distribution]

---
# Choice of Predictors

One challenge remaining for the researcher is the .hi-medgrn[definition of predictors]

--

  * Which predictors to use 
--

  * Which years to match on
--
  
[Ferman, Pinto, and Possebom (2020)](https://onlinelibrary.wiley.com/doi/abs/10.1002/pam.22206) go into great detail regarding how to properly select specifications of synthetic controls. Their punchline:
--

* Models including more pre-treatment outcome lags as predictors are better at controlling for unobserved confounders
--

* The possibilities for "specification searching" are higher with more pre-treatment periods used for matching
--

* .hi-blue[Best:] present multiple results under common specifications
  * If the result is robust to these different predictor choices, then the preferred specification isn't cherry-picked!

---
# Table of Contents


1. [Prologue](#prologue)

3. [Matching](#match)

4. [Canonical Synthetic Control](#canon)




```{r gen_pdf, include = FALSE, cache = FALSE, eval = FALSE}
infile = list.files(pattern = 'Pt1.html')
pagedown::chrome_print(input = infile, timeout = 200)
```